---
title: "GWAS_rrBLUP"
author: "Leah Treffer, adapted from Jared Crain"
output: html_document
editor_options: 
  chunk_output_type: console
bibliography: citations.bib
nocite: |
  @*
---

# Objective

1. Run a GWAS analysis for measured traits 
	- C7 Salina 2018, 2019, 2020
	- C10 Salina 2021, 2022
	- C10 Olathe 2021, 2022
2. Compare results to those from 1.1_GWAS_FarmCPU

```{r setup, eval=TRUE, echo= FALSE, include=FALSE}
rm(list=ls(all=TRUE))
#set current working directory
#setwd()
setwd('/Volumes/Backup_Plus/AM/New_GWAS/2023.01.27/IWG_GWAS')
```

Packages used:

```{r library, eval=TRUE, echo= TRUE, include=FALSE}
require(reshape2)
require(rrBLUP)
require(factoextra)
library(tidyr)
library(stringr)
library(dplyr)
library(tibble)
install.packages("devtools")
library("devtools")
install_github("jdstorey/qvalue")
require(qvalue)
require(qqman)
#install.packages("errorist")
#library(errorist)
```

Custom function from Jared Crain:

```{r get_functions, eval=TRUE, echo= TRUE, include=FALSE}
source('./scripts/Function_R.r') #source functions

#function to get value to make FDR line modified from rrBLUP GWAS function
fdr_line <- function(q.ans = 'list of qvulues', logp = 'list of -log(P) values', fdr.level = 'Level for False Discovery Rate'){ #modified code from GWAS function of rrBLUP
  fdr_line <- NULL #set up fdr line
  temp <- cbind(q.ans, logp) #organizes data similar to rrBLUP
  temp <- temp[order(temp[, 1]), ] #order data
  #######from GWAS function (EXACT) of rrBLUP #cite rrBLUP
  if (temp[1, 1] < fdr.level) {
    temp2 <- tapply(temp[, 2], temp[, 1], mean) #averages over qvalues of the same value
    qvals <- as.numeric(rownames(temp2)) #extract all unique qvalues
    x <- which.min(abs(qvals - fdr.level)) #finds qvalue - fdr that has lowest value
    first <- max(1, x - 2)
    last <- min(x + 2, length(qvals))
    if ((last - first) < 4) {
      last <- first + 3
    }
    splin <- smooth.spline(x = qvals[first:last], y = temp2[first:last], 
                           df = 3) #run spline to find smoothed line
    ##Modified from GWAS function of rrBLUP to return FDR value for fdr level
    fdr_line <- predict(splin, x = fdr.level)$y #get value for FDR line
                                  
  }
 return(fdr_line) #return value for FDR line
}
```

```{r data, eval=TRUE, echo= FALSE, include=FALSE}
GENO <- read.delim("data/Intermediate_Files/C10_OLA_2022_numeric.txt", head=TRUE) #load genotype calls formated for GAPIT
geno <- GENO #make copy to work with
rownames(geno) <- geno[,1] # column to rownames
geno <- geno[,-1] #remove first column 
geno <- geno - 1 #change geno numeric format from 012 to -101

pheno <- read.table("data/Intermediate_Files/C10_OLA_2022_pheno.txt", head=TRUE)
```

# Set up rrBLUP GWAS

**Population structure** : We will first get a realized relationship matrix by calculating the additive relationship matrix. Then, principle comonenet analysis can be done for the genotype matrix followed by evaluting population structure through eigen values taken from the realized relationship matrix. The output plot is created using this equation: $lambda/sum(lambda)$ where lambda is the variance explained. 

```{r population_structure, eval=TRUE, echo=FALSE, include=TRUE}

#get realized relationship matrix
# A.mat() Calculates the realized additive relationship matrix
A <- A.mat(geno) #realized relationship matrix (G matrix)

#PCA on genotype matrix
prmarker <- prcomp(geno)
fviz_eig(prmarker)
#first 2 components

#look for population structure
eig.result <- eigen(A) #get Eigen values from the A matrix
lambda <- eig.result$values #extract variance explained
  
pdf(file = './output/rrBLUP/C10_OLA_2022/AM_Genotypes_Structure.pdf', height = 6, width = 6) #save plot
plot(lambda/sum(lambda), ylab = 'Fraction Explained Num Offspring', main = 'Fraction explained by Genotypes') #plot fraction explained for each value
dev.off()

```

Our plot of population structure shows a sharp decline at 2 principle components. A change in slope is a guideline for deciding how many principle components to use. There is very little population structure, which is in line with other analysis of *Thinopyrum intermedium* [@citation_here_2004]. 

```{r gwas_rrBLUP, eval=FALSE, echo=FALSE, include=FALSE}
#remove unstructured chromosomes from mapping
geno <- geno[, !grepl('SUN', colnames(geno), ignore.case = TRUE) ] 

#GWAS fit geno file for gwas
rr_gwas <- t(geno) #transpose

#format marker names
markers <- as.data.frame(rownames(rr_gwas)) #extract maker names
colnames(markers)[1] <- 'marker_name'
markers$pos <- gsub('^.*_', '', markers$marker_name) #split string and extract marker position
markers$pos <- as.numeric(markers$pos) #convert to numeric
markers$chrom <- gsub('_.*$', '', markers$marker_name) #split string and extract chromosome
markers$chrom <- sub('S', '', markers$chrom) #remove string 
markers <- markers[order(markers$chrom, markers$pos),] #orders markers by chromosome and then position
markers$marker <- 1:nrow(markers) #add marker name as numeric
markers <- markers[ , c(4:1)] #rearrange columns

#merge markers with gwas to format for rrBLUP
rr_gwas <- as.data.frame(rr_gwas) #convert to data frame
rr_gwas$marker_name <- rownames(rr_gwas) #move marker_names to rownames
rownames(rr_gwas) <- NULL #remove rownames

rr_gwas <- merge(markers, rr_gwas, by = 'marker_name') #join marker number, name, and chromosome for rrBLUP

rr_gwas <- rr_gwas[ order(rr_gwas$marker), ] #sort marker order
rr_gwas <- rr_gwas[ , -1] #remove first column

all(geno$names == colnames(rr_gwas)[4:ncol(rr_gwas)]) #check that they match

#loop over all columns
for(i in 2:ncol(pheno)){ #set up forloop

#run GWAS
  #Performs genome-wide association analysis based on the mixed model (Yu et al. 2006)
ans.emmax <- GWAS(pheno = pheno[, c(1, i)], geno = rr_gwas, P3D = TRUE, n.core = 1 , K = A, n.PC = 2, plot = TRUE) 
  #PD3 (population parameters previously determined)
  #PD3 = FALSE  EMMA with REML (Kang et al. 2008)
  #PD3 = TRUE  EMMAX (Kang et al. 2010); faster but can underestimate significance compared to P3D=FALSE

#save results and plots
saveRDS(ans.emmax, file = paste('./output/rrBLUP/C10_OLA_2022/', colnames(pheno)[i], '.RDS', sep = '')) 
#plotting with qqman #tutorial ran with 2 PC 
  #data frame where the first three columns are the marker name, chromosome, and position, and subsequent columns are the marker scores (−log10p) for the traits
}

#clean up 
rm(ans.emmax, A, eig.result, geno, markers, pheno, prmarker, rr_gwas, lambda)
```

```{r get_results, eval=TRUE, echo=FALSE, include=FALSE}

gwas <- list.files(path = './output/rrBLUP/C10_OLA_2022/', pattern = '*.RDS')

result <- NULL #start dataframe for results

for(i in 1:length(gwas)){ #loop over GWAS and load results
  trait <- gwas[i] #get trait_id
  
  values <- readRDS(paste('./output/rrBLUP/C10_OLA_2022/', trait, sep = '')) #load file
  values$trait_id <- colnames(values)[4] #get column name
  colnames(values)[4] <- 'LOD' #set standard LOD column name
  values$trait_id <- gsub('^X', '', values$trait_id)
  result <- rbind(result, values[, 2:5]) #write out results
}

result <- dcast(result, formula = chrom + pos ~ trait_id, value.var = 'LOD', fun.aggregate = mean) #write out LOD scores

#write results
write.csv(result, file = './output/rrBLUP/C10_OLA_2022/SPAM_AM_GWAS_Results.csv', row.names = FALSE, quote = FALSE) 

#clean up
rm(gwas, i, trait, values)

```

```{r graph_results, eval=TRUE, echo=FALSE, include=TRUE}

traits <- colnames(result)[3:ncol(result)] #get trait_ids
traits <- unique(traits) #make sure trait names are unique 

#for each trait loop through and create graphs for each year and trial as applicable
for(i in 1:length(traits)){ #loop over traits 
  
  if(paste(traits[i]) %in% colnames(result) ){ #if exist prepare for graph
    am_C <- result[, c('chrom', 'pos', paste(traits[i])), ] #table of chrom, pos, and one trait
    colnames(am_C)[3] <- 'LOD' #set LOD score, change column name from trait id to 'LOD'
    am_C$pval <- 10^-am_C$LOD #make p value instead -log(pvalue)
    am_C_gw_fdr <- fdr_line(q.ans = qvalue(am_C$pval)$qvalues, logp = am_C$LOD, fdr.level = 0.05)  #get genome wide fdr 
  }
  
  } #close for loop


key <- as.data.frame(unique(result$chrom)) #get unique chromosomes
key$number <- 1:nrow(key)
colnames(key)[1] <- 'chrom' #rename column

result <- merge(result, key, by = 'chrom')  #add chromosome number to results 
result <- result[, c(1:2, 9, 3:8)] #rearrange columns *always check this to have all columns 


#get pvalues
for(i in 4:ncol(result)){ #loop over columns and make pvalues
  x <- result[, c(1:3, i)] #columns before traits
  trait_id <- colnames(x)[4]
  colnames(x)[4] <- 'trait'
  x$pval <- 10^-x$trait #make p value instead -log(pvalue)

  #get genome wide fdr
gw_fdr <- fdr_line(q.ans = qvalue(x$pval)$qvalues, logp = x$trait, fdr.level = 0.05)
  pdf(file = paste('./output/rrBLUP/C10_OLA_2022/QQ_Plot_', trait_id, '.pdf', sep = ''), height = 4, width = 4)
qq( x$trait)#make a qq plot
dev.off()

#rename columns
colnames(x)[c(1, 3, 2)] <- c('SNP', 'CHR', 'BP') 
x <- x[order(x$CHR, x$BP), ] #order markers

pdf(file = paste('./output/rrBLUP/C10_OLA_2022/Manhattan_', trait_id, '.pdf', sep = ''), height = 6, width = 9 ) #save as a smaller pdf


#Model AM.
if(is.null(gw_fdr)){
  gw_fdr <- FALSE
}else{
  gw_fdr <- gw_fdr
}

manhattan(x, main=trait_id, p = 'pval', suggestiveline = FALSE, col = c('darkseagreen2', 'darkseagreen'), ylim = c(0, 8), genomewideline = gw_fdr) #make main plot
#chr:numeric string denoting the column name for the chromosome#bp:numeric string denoting the column name for the chromosomal position#p:numeric string denoting the column name for the p-value#snp:character string denoting the column name for the SNP name (rs number)
dev.off()

  
  }


#clean up
rm(am_C,key,x,am_C_gw_fdr,gw_fdr,i,trait_id)
```

# Significant markers

```{r get_significant_markers, eval=TRUE, echo=FALSE, include=TRUE}

sig_results <- result[,c('chrom','pos','number')]
sig_results$snp <- paste0(sig_results$chrom,'_',sig_results$pos)
result$snp <- paste0(result$chrom,'_',result$pos)

for(i in traits){ #loop over columns and make pvalues
  temp <- result[,c('chrom','pos','number','snp',i)]
  x <- result[,c('chrom','pos','number','snp',i)]
    trait_id <- colnames(x)[5]
    colnames(x)[5] <- 'trait'
    x$pval <- 10^-x$trait #make p value instead -log(pvalue)
    gw_fdr <- fdr_line(q.ans = qvalue(x$pval)$qvalues, logp = x$trait, fdr.level = 0.05) #get genome wide fdr
    
  sig <- temp[temp[[i]] > gw_fdr,] #LOD values above the fdr line for that trait
  sig_results <- merge(sig_results, sig[,c(4,5)], by="snp", all=TRUE)
}

df <- filter(sig_results, rowSums(is.na(sig_results[,c(5:ncol(sig_results))])) != ncol(sig_results[,c(5:ncol(sig_results))])) #remove all rows where all trait columns are NA

write.csv(df, file = './output/rrBLUP/C10_OLA_2022/Significant_Markers_C10_OLA_2022.csv', row.names = FALSE, quote = FALSE) #write results for BLAST or gene search

#clean up
rm(GENO,df,result,sig,sig_results,temp,x,gw_fdr,i,trait_id,traits)
```

