---
title: "2.6_Ref+MAF+AlleleFrequencies_2025.Rmd"
author: "Leah Treffer"
date: "2025-12-10"
output: html_document
---

```{r}
library(here)
knitr::opts_knit$set(root.dir = here::here())
here::i_am("2.6_Ref+MAF+AlleleFrequencies_2025.Rmd")
```

```{r include=FALSE}
require(reshape2)
library(tidyr)
library(stringr)
library(dplyr)
library(data.table)
install.packages("gplots")
install.packages("scatterplot3d") #The downloaded link at: http://cran.r-project.org/package=scatterplot3d
library(bigmemory)
```

Look at only traits with significant SNPs and add ref+alleles to these trait tables

Reference allele was computed using Bedtools on beocat: 
% module load BEDTools #load module BEDTools/2.29.1-GCC-8.3.0 
% bedtools getfasta -fi /homes/jcrain/IWG/Thinopyrum_intermedium_V3_release/Thinopyrum_intermedium/sequences/Thinopyrum_intermedium.fasta -bed BedTools_Reference.bed -tab -name -fo FHB_Reference_Allele.bed 

```{r}
#Add Ref and Alleles to table
refall <- read.table(here("data","Original_Files","FHB_Reference_Allele.bed"),header = FALSE, sep="\t",stringsAsFactors=FALSE, quote="") #load reference list 
colnames(refall)[1] <- "rs" #rename snp column to rs to match hapmap format

hapmap <- read.table(here('data','Original_Files','FHB_Imputed.hmp.txt'), header = TRUE, check.names = FALSE)
hm <- hapmap[,c(1:4)] #get snp block  
colnames(hm)[1] <- "SNP"

rs <- as.data.frame(gsub("\\:.*","",refall$rs)) #remove characters after the snp: section 
colnames(rs)[1] <- "SNP" #rename snp column to rs to match hapmap format
rs <- cbind(refall, rs) #combine the snp block to the map block
rs <- rs[,c(3,2)] #reorder
colnames(rs)[2] <- "Ref" #rename 
hmm <- merge(rs, hm, by="SNP", all=TRUE) #merge files

rm(hm)
```

Add Reference and Alt alleles to sig loci list

```{r separatePheno}
#Use list of Significant SNPs
#get traits that had significant p.values
trait <- read.csv(here('data','Intermediate_Files','SigLocilist_2025.csv'), head=TRUE)

# merge
addAllele <- merge(trait, hmm[,c(1:3)], by="SNP")
```

Use geno files and Ref Allele list retrieved from jcrain beocat to determine which allele is major and minor in numeric format geno file 

```{r}
geno <- read.table(here(file = 'data','Original_Files','FHB_for_GS_GWAS.txt'), header = TRUE, row.names = 1, stringsAsFactors = FALSE) #load genotype calls, slow

plants <- read.csv(here(file = 'data','Original_Files','plants.csv')) #load ID list
plantC07 <- plants[str_detect((plants$source_id), "17SGH"),] #keep only data from one population
plantC10 <- plants[str_detect((plants$source_id), "20SGH"),] #keep only data from one population
plantC10SAL <- plants[str_detect((plants$experiment_id), "20_SAL"),] #keep only data from one site
plantC10SAL <- plantC10SAL[str_detect((plantC10SAL$source_id), "20SGH"),] #keep only data from one site
plantC10OLA <- plantC10[str_detect((plantC10$experiment_id), "20_OLA"),] #keep only data from one site
```

```{r}
#Reference Alleles
#make table of snp, ref, and alleles 
rss <- hmm[,c(1:3)]

#Hapmap remove unnecessary columns, get just C7s
hapm <- hapmap[,c(1,12:ncol(hapmap))]
hh <- hapm[,c(2:ncol(hapm))]
a <- hapm[,c(1)]

#get map block for just C7 population
hh07 <- hh[ , grep("17SGH", colnames(hh))] 
#get map block for just C10 population 
pSAL <- c(na.omit(unique(plantC10SAL$source_id))) #in C10 pop, get only one site
names.useS <- names(hh)[(names(hh) %in% pSAL)] # make list of names from p that are in hh
hh10S <- hh[, names.useS] #subset
pOLA <- c(na.omit(unique(plantC10OLA$source_id))) #in C10 pop, get only one site
names.useO <- names(hh)[(names(hh) %in% pOLA)] # make list of names from p that are in hh
hh10O <- hh[, names.useO] #subset

hpmp07 <- cbind(a, hh07) #combine the snp block to the map block
rownames(hpmp07) <- hpmp07$a #make first column into row names
hpmp07 <- hpmp07[,-1] #remove first column 

hpmp10S <- cbind(a, hh10S) #combine the snp block to the map block
rownames(hpmp10S) <- hpmp10S$a #make first column into row names
hpmp10S <- hpmp10S[,-1] #remove first column 
hpmp10O <- cbind(a, hh10O) #combine the snp block to the map block
rownames(hpmp10O) <- hpmp10O$a #make first column into row names
hpmp10O <- hpmp10O[,-1] #remove first column 

#Get hapmap in same orientation as geno file
#get just the SNPs that are in our QTL set, this will reduce the size of files 
trait7 <- trait[str_detect((trait$plot),"C07"),]
trait10S <- trait[str_detect((trait$plot), "C10_SAL"),]
trait10O <- trait[str_detect((trait$plot), "C10_OLA"),]
list7 <- unique(trait7$SNP) #remove duplicate snps so each is unique 
list10S <- unique(trait10S$SNP) #remove duplicate snps so each is unique 
list10O <- unique(trait10O$SNP) #remove duplicate snps so each is unique 

#C07 genotype list
g <- as.data.frame(rownames(geno))
#C10 genotype lists
g10S <- geno[rownames(geno) %in% pSAL, ]
g10S <- as.data.frame(rownames(g10S))
g10O <- geno[rownames(geno) %in% pOLA, ]
g10O <- as.data.frame(rownames(g10O))

colnames(g)[1] <- "taxa"
colnames(g10S)[1] <- "taxa"
colnames(g10O)[1] <- "taxa"


#C07
#loop through each significant snp that is in our QTL list into it's own row, transpose, and put table back together by merging 
for (j in list7){ 
  row <- hpmp07[j,]
  data2 <- tibble::rownames_to_column(row, "row_names") #convert row names into first column of data
  data2 <- rbind(colnames(data2), data2) #make copy of header columns and adds them as the first row
  colnames(data2) <- paste0('V', seq_len(ncol(data2))) #changes header to 'V1..'
  data2 <- as.data.frame(t(data2)) # flip table 
  colnames(data2) <- data2[1, ] #make first row into header
  data2 = data2[-1,] #remove first row
  colnames(data2)[1] <- "taxa" 
  g <- merge(data2, g, by="taxa", all=TRUE) #merge files
}

#clean up
rm(data2, row)

#C10 SAL
#loop each snp that is in our QTL list into it's own row, transpose, and put table back together by merging 
for (j in list10S){ 
  row <- hpmp10S[j,]
  data2 <- tibble::rownames_to_column(row, "row_names") #convert row names into first column of data
  data2 <- rbind(colnames(data2), data2) #make copy of header columns and adds them as the first row
  colnames(data2) <- paste0('V', seq_len(ncol(data2))) #changes header to 'V1..'
  data2 <- as.data.frame(t(data2)) # flip table 
  colnames(data2) <- data2[1, ] #make first row into header
  data2 = data2[-1,] #remove first row
  colnames(data2)[1] <- "taxa" 
  g10S <- merge(data2, g10S, by="taxa", all=TRUE) #merge files
}

#clean up
rm(data2, row)

#C10 OLA
#loop each snp that is in our QTL list into it's own row, transpose, and put table back together by merging 
for (j in list10O){ 
  row <- hpmp10O[j,]
  data2 <- tibble::rownames_to_column(row, "row_names") #convert row names into first column of data
  data2 <- rbind(colnames(data2), data2) #make copy of header columns and adds them as the first row
  colnames(data2) <- paste0('V', seq_len(ncol(data2))) #changes header to 'V1..'
  data2 <- as.data.frame(t(data2)) # flip table 
  colnames(data2) <- data2[1, ] #make first row into header
  data2 = data2[-1,] #remove first row
  colnames(data2)[1] <- "taxa" 
  g10O <- merge(data2, g10O, by="taxa", all=TRUE) #merge files
}

#clean up
rm(data2, row)
```

```{r convert}
#convert coded hapmap format to specific allele
#Genotype:Code
## AA:A
## CC:C
## GG:G
## TT:T
## AG:R
## CT:Y
## CG:S
## AT:W
## GT:K
## AC:M
## missing:N

#replace one letter code with the genotype 
g[g == "A"] <- "AA" #replace
g[g == "C"] <- "CC" #replace
g[g == "G"] <- "GG" #replace
g[g == "T"] <- "TT" #replace
g[g == "R"] <- "AG" #replace
g[g == "Y"] <- "CT" #replace
g[g == "S"] <- "CG" #replace
g[g == "W"] <- "AT" #replace
g[g == "K"] <- "GT" #replace
g[g == "M"] <- "AC" #replace
g[g == "N"] <- "NN" #replace
#replace one letter code with the genotype 
g10S[g10S == "A"] <- "AA" #replace
g10S[g10S == "C"] <- "CC" #replace
g10S[g10S == "G"] <- "GG" #replace
g10S[g10S == "T"] <- "TT" #replace
g10S[g10S == "R"] <- "AG" #replace
g10S[g10S == "Y"] <- "CT" #replace
g10S[g10S == "S"] <- "CG" #replace
g10S[g10S == "W"] <- "AT" #replace
g10S[g10S == "K"] <- "GT" #replace
g10S[g10S == "M"] <- "AC" #replace
g10S[g10S == "N"] <- "NN" #replace
#replace one letter code with the genotype 
g10O[g10O == "A"] <- "AA" #replace
g10O[g10O == "C"] <- "CC" #replace
g10O[g10O == "G"] <- "GG" #replace
g10O[g10O == "T"] <- "TT" #replace
g10O[g10O == "R"] <- "AG" #replace
g10O[g10O == "Y"] <- "CT" #replace
g10O[g10O == "S"] <- "CG" #replace
g10O[g10O == "W"] <- "AT" #replace
g10O[g10O == "K"] <- "GT" #replace
g10O[g10O == "M"] <- "AC" #replace
g10O[g10O == "N"] <- "NN" #replace
```

```{r}
#find what allele is major for each SNP

# add populations to a list
test <- list(g, g10S, g10O)
names(test) <- c('C07', 'C10SAL', 'C10OLA') #rename list

# loop through each population to find major allele based on allele summaries
for(k in names(test)){
  count <- test[[k]] #make copy of test to work with 
  count <- as.data.frame(count)
  rownames(count) <- count$taxa #make first column into row names
  count <- count[,-1] #remove first column

  #Table with count of each SNP's allele summary 
  # compute unique levels in data frame
  lvls <- unique(unlist(count))
  # apply the summation per value 
  freq <- sapply(count, 
                 function(x) table(factor(x, levels = lvls, 
                                         ordered = TRUE)))

  #Table of SNPs, Reference, Alleles, and Frequencies 
  freq2 <- as.data.frame(t(freq)) #transpose and change to dataframe
  freq2 <- tibble::rownames_to_column(freq2, "rs") #convert row names into first column of data
  colnames(rss)[1] <- "rs"
  freq2 <- merge(freq2, rss, by="rs") #merge to add in ref allele 
  freq2 <- freq2[,c(1,12,13,2:11)] #reorder

  freq2$total <- rowSums(freq2[,c(4:13)]) #Total taxa with genotypes

  #Genotype Frequency = AA / Total = (# of individuals with genotype) / (total individuals)
  Genotypefrequency <- cbind(freq2,
    fCT <- ((freq2$CT) / (freq2$total)),
    fCC <- ((freq2$CC) / (freq2$total)),
    fTT <- ((freq2$TT) / (freq2$total)),
    fAT <- ((freq2$AT) / (freq2$total)),
    fAA <- ((freq2$AA) / (freq2$total)),
    fAG <- ((freq2$AG) / (freq2$total)),
    fGG <- ((freq2$GG) / (freq2$total)),
    fCG <- ((freq2$CG) / (freq2$total)),
    fAC <- ((freq2$AC) / (freq2$total)),
    fGT <- ((freq2$GT) / (freq2$total))
  )
  #change column names
  colnames(Genotypefrequency)[15] <- "fCT"
  colnames(Genotypefrequency)[16] <- "fCC"
  colnames(Genotypefrequency)[17] <- "fTT"
  colnames(Genotypefrequency)[18] <- "fAT"
  colnames(Genotypefrequency)[19] <- "fAA"
  colnames(Genotypefrequency)[20] <- "fAG"
  colnames(Genotypefrequency)[21] <- "fGG"
  colnames(Genotypefrequency)[22] <- "fCG"
  colnames(Genotypefrequency)[23] <- "fAC"
  colnames(Genotypefrequency)[24] <- "fGT"

  #Allelic Frequency (AF)
  #GF= genotype frequencies
  GF <- Genotypefrequency #create copy with shorter name
  #Allelic Frequency = genomic frequency of homozygous + 1/2 genomic frequency heterozygous for specific Base  
  #AF= GF(XX) + 1/2GF(Xx) 
  MAfrequency <- cbind(freq2,
  fA <- ((GF$fAA + (GF$fAT / 2) + (GF$fAC / 2) + (GF$fAG / 2))),
  fT <- ((GF$fTT + (GF$fAT / 2) + (GF$fCT / 2) + (GF$fGT / 2))),
  fG <- ((GF$fGG + (GF$fAG / 2) + (GF$fCG / 2) + (GF$fGT / 2))),
  fC <- ((GF$fCC + (GF$fCT / 2) + (GF$fAC / 2) + (GF$fCG / 2)))
  )
  #change column names
  colnames(MAfrequency)[15] <- "A"
  colnames(MAfrequency)[16] <- "T"
  colnames(MAfrequency)[17] <- "G"
  colnames(MAfrequency)[18] <- "C"

  freqs <- merge(Genotypefrequency,MAfrequency[,c(1,15:18)], by="rs")
  colnames(freqs)[1] <- "SNP"
  colnames(freqs)[15] <- "genotypic_frequency CT"
  colnames(freqs)[16] <- "genotypic_frequency CC"
  colnames(freqs)[17] <- "genotypic_frequency TT"
  colnames(freqs)[18] <- "genotypic_frequency AT"
  colnames(freqs)[19] <- "genotypic_frequency AA"
  colnames(freqs)[20] <- "genotypic_frequency AG"
  colnames(freqs)[21] <- "genotypic_frequency GG"
  colnames(freqs)[22] <- "genotypic_frequency CG"
  colnames(freqs)[23] <- "genotypic_frequency AC"
  colnames(freqs)[24] <- "genotypic_frequency GT"
  colnames(freqs)[25] <- "allelic_frequency A"
  colnames(freqs)[26] <- "allelic_frequency T"
  colnames(freqs)[27] <- "allelic_frequency G"
  colnames(freqs)[28] <- "allelic_frequency C"

  name_freqs <- paste0(k,"_Frequencies")
  
  assign(name_freqs, freqs) # save
}

rm(count,freq,freq2,freqs,Genotypefrequency,GF,MAfrequency,lvls,j,fA,fAA,fAC,fAG,fAT,fC,fCC,fCG,fCT,fG,fGG,fGT,fT,fTT,test,rss,hh)
```

```{r}
# output is C07_Freqnecies, C10SAL_Frequencies, C10OLA_Frequencies
# how to save? 
Frequency_table <- rbind(C07_Frequencies,C10OLA_Frequencies,C10SAL_Frequencies) #219 rows
Ref_MAF_Freq <- merge(addAllele, Frequency_table[,c("SNP","GG","AG","AA","CG","CC","TT","GT","CT","AT","AC","genotypic_frequency CT","genotypic_frequency CC","genotypic_frequency TT","genotypic_frequency AT","genotypic_frequency AA","genotypic_frequency AG","genotypic_frequency GG","genotypic_frequency CG","genotypic_frequency AC","genotypic_frequency GT","allelic_frequency A","allelic_frequency T","allelic_frequency G","allelic_frequency C")], by="SNP")
# merged table has 228 obs but the significant loci list has 226. Issue is SV02_452036252 has duplicate rows and two different frequencies for the same trait-plot. In the individual cycle's frequency tables, alleles have different frequencies between olathe and salina locations indicating the genotypes present at the two locations have different allele frequencies. 
# issues is that the merge dosn't know this and copies both frequencies lists into the SNPS instead of matching to the specific trait-plot. 
Ref_MAF_Freq <- Ref_MAF_Freq[-c(183,184),]
```

```{r}
# save
# write.csv(Ref_MAF_Freq, here('data','Intermediate_Files','SigLociAlleles_2025.csv'))
```

